---
title: "Arsen_ExampleRCode"
author: "Celia Arsen"
date: "March 4, 2019"
output: html_document

Data: Provided by Professor Daniel Hamermesh of the University of Texas at
Austin and were used in his paper with Amy Parker, “Beauty in the Classroom: Instructors’
Pulchritude and Putative Pedagogical Productivity,” Economics of Education Review, August
2005, Vol. 24, No. 4, pp. 369-376. 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Import teacher ratings dataset
```{r}
setwd("C:/Users/Celia/OneDrive - barnard.edu/Documents/Computer Science/ExampleCode")
teacher_ratings <- read.csv("evaluation_data.csv", header = TRUE)
```

Look at some summary statistics of the dataset
```{r}
#variable names
names(teacher_ratings)
#look at rows 3 through 10 for variable 'course_eval'
teacher_ratings$course_eval[3:10]
#summary statistics
mean(teacher_ratings$course_eval[teacher_ratings$female==0])
mean(teacher_ratings$course_eval[teacher_ratings$female==1])
```


```{r}
#add labels to data 
teacher_ratings$female <- factor(teacher_ratings$female,
  levels = c(0,1),
  labels = c("male", "female"))
teacher_ratings$intro <- factor(teacher_ratings$intro,
   levels = c(0,1),
   labels = c("non-intro", "intro"))

#create a cross-tab to look at distribution of female instructors and introductory classes
female_intro <- table(teacher_ratings$female, teacher_ratings$intro, dnn = c("gender", ""))
female_intro

#What proportion of women are teaching intro classes
prop.table(female_intro, 1)
```

Is there a visible relationship between beauty ratings and course evaluations?
```{r}
#libarary(ggplot)

eval.beauty.plot <- ggplot(teacher_ratings, aes(x=beauty, y=course_eval)) +
  geom_point(alpha=.5) + 
  geom_smooth(method='lm',formula=y~x) +
  xlab("Beauty Rating") + 
  ylab("Course Evaluation Rating")
                    
eval.beauty.plot
```

We can see that there is a slightly positive relationship between beauty and course evaluation ratings. 
That is, teachers who had higher beauty ratings tended to have higher course evaulation ratings. 
However, there is a lot of variance. Is this relationship statistically significant?
```{r}
cor.test(teacher_ratings$course_eval, teacher_ratings$beauty, use="complete.obs", method="pearson")
```

Bar chart comparing the average course evaluation rating between men and women
```{r}
eval.beauty.plot <- ggplot(teacher_ratings, aes(x=female, y=course_eval)) +
  geom_bar(stat="summary", fun.y = "mean", width = 0.5, fill = "lavender", colour = "white") +
  xlab("Gender") + 
  ylab("Course Evaluation Rating")

eval.beauty.plot
```

Are differences in male and female beauty ratings significant?
```{r}
t.test(teacher_ratings$course_eval ~ teacher_ratings$female)

```

Test if beauty has a significantly larger impact on course evaluations for women than men
Dependent variable is course_eval
Independent variables (with interaction) are female and beauty
```{r}
beauty_female_fit <- lm(course_eval ~ female*beauty, data=teacher_ratings) 
summary(beauty_female_fit)

```


